Supplemental Information:\\ Harnessing Design Principles from Glass Sponges for Structurally Robust Lattice Structures

### 2018-09-05 Supplemental-FernandesWeaverBertoldi_AdvMaterials_2018.tex
Updates from ShareLaTeX
Before Katia begins to edit

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
\newcommand{\MF}[1]{\noindent\color{ForestGreen}#1\normalcolor}

@@

\begin{document}
\maketitle
% \doublespace
% \linenumbers
\section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
The periodic structures investigated in this study are inspired by the skeleton of the hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket". In this section we provide a detailed description of the sponge geometry and measured dimensions.

@@

Finally, it should also be noted that the sponge is reinforced by external ridges that extend perpendicular to the surface of the cylinder and spiral the cage at an angle of $45^\text{o}$. However, in this paper we do not report the effects of these ridges on it's structural performance %and treat the periodic structure as a planar (no out of plane deformation) unit cell.
% \figll{SpongeCylinder}{0.35}{Close-up photograph of Hexactinellid sponge Euplectella aspergillum.}
\figll{Sponge}{0.9}{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.  (c)  Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure. (e) Unit cell of the idealized model.}

\section{Designs Considered}
In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only. In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations the total amount of material (i.e. volume or mass) is conserved. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

%In this study, we consider four different planar periodic lattice structures and compare their performance in terms of buckling resistance and stiffness. All structures are made of struts with circular cross section and are assumed to only deform in-plane.  In an effort to provide a fair performance comparison between all four designs, we scale the different diameters of each design to maintain a constant total volume for each unit cell. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

\subsection{Polygonal Cross-section}
\subsubsection{Design A}
From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ . Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
\begin{equation}
	V_{A,nd}=8LD_{A,nd}t
\end{equation}
\begin{equation}
	V_{A,d}=8\sqrt{2}LD_{A,d}t
\end{equation}
Let us enter this into our biological observation, which yields:
\begin{equation}
	{V_{A,nd}\over V_{A,d}}\approx 1.4\approx\sqrt{2}={8LD_{A,nd}t\over8\sqrt{2}LD_{A,d}t}
\end{equation}
Therefore, we can simplify this relationship as
\begin{equation}
	D_{A,nd}=2D_{A,d}
\end{equation}

\subsubsection{Design B}
For this design the volume of non-diagonal elements remain the same, namely,
\begin{equation}
		V_{B,nd}=8LD_{B,nd}t.
\end{equation}
However, the volume for the diagonal elements will is different as a result the change in total length, namely,
\begin{equation}
		V_{B,d}=4\sqrt{2}LD_{A,d}t.
\end{equation}
In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
\begin{equation}
		{V_{B,nd}\over V_{B,d}}\approx 1.4\approx\sqrt{2}={8LD_{B,nd}t\over4\sqrt{2}LD_{B,d}t}
\end{equation}
which simplifies to
\begin{equation}
	D_{B,nd}=D_{B,d}
\end{equation}
and
\begin{equation}
	D_{B,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design D}
For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements. Thus we can formulate the total volume of Design A as
\begin{equation}
	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
\end{equation}
Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
\begin{equation}
V_{A,nd}+V_{A,d}=\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}
\end{equation}

We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
\begin{equation}
\left(1+{1\over\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
\end{equation}
which yields a relationship between the thicknesses of the non diagonal elements as
\begin{equation}
	D_{D,nd}=\left(1+{1\over\sqrt{2}}\right)D_{A,nd}
\end{equation}

\subsection{Cylindrical Cross-section}

\subsubsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{TwoDiag}). If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
respectively. Note that in this study we use 
\begin{equation}
\frac{D_{A,nd}}{L}=0.1,
\end{equation}
since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

@@

\figll{TwoDiag}{0.45}{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system. The horizontal and vertical struts have length $L$ and circular cross section with diameter $D_{A,nd}$ and, as with the sponge, we assume $D_{A,nd}/L=0.1$. The diagonal elements have a circular cross section  with diameter $D_{A,d}=2 D_{A,nd}$.}

\subsubsection{Design B}
Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{OneDiag}). However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. It follows that the non-diagonal and diagonal volumes are given by
\begin{equation}
V_{B,nd}=V_{A,nd}=2\pi L D_{B,nd}^2
\end{equation}
and
\begin{equation}
V_{B,d}=2\sqrt{8} L \left(\pi \frac{{D}_{B,d}^2}{4}\right),
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2}, as well as the above volumes, we  obtain 
\begin{equation}
{{D}_{B,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{B,d}}{{D}_{B,nd}}=\frac{1}{\sqrt{2}}.
\end{equation}

 \figll{OneDiag}{0.45}{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells. However, instead of having two  diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{B,nd}$. The diagonal elements have a circular cross section  with diameter $D_{B,d}={D_{B,nd}}/{\sqrt{2}}$.}

\subsubsection{Design C}
Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{FullDiag}). For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
\begin{equation}
V_{C,nd}=V_{A,nd}=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}
V_{C,d}=V_{A,d}=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2} we  obtain 
\begin{equation}
{{D}_{C,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{C,d}}{{D}_{C,nd}}=\frac{1}{2}.
\end{equation}

\figll{FullDiag}{0.45}{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{C,nd}$. The diagonal elements have a circular cross section  with diameter $D_{C,d}={D_{C,nd}}/{2}$.}

\subsubsection{Design D} 
Design D comprises only the square grid without diagonal reinforcement (\cref{NoDiag}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}. As such, for this design we allocate the total material volume to the non-diagonal elements. Since
\begin{equation}
V_{D,T}=V_{D,nd}=V_{A,nd}=2\pi L D_{D,nd}^2,
\end{equation}
using the constraint provided by \cref{con1} we obtain
\begin{equation}
{D_{D,nd}}={D_{A,nd}}\sqrt{1+\frac{\sqrt{2}}{4}}. 
\end{equation}

\figll{NoDiag}{0.45}{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{D,nd}$.}

>>>>

\newcommand{\KB}[1]{\noindent\color{blue}$\Longrightarrow$ #1\normalcolor}
\newcommand{\mf}[1]{\colorbox{blue!10}{\color{color3}#1}}

@@

\begin{document}
\maketitle
% \doublespace
% \linenumbers
\section{Structure of the Hexactinellid sponge \textit{Euplectella aspergillum}}\label{sec:params}
The periodic structures investigated in this study are inspired by the skeleton of the Hexactinellid sponge \textit{Euplectella aspergillum (sp.)}, commonly known as the "Venus' flower basket". In this section we provide a detailed description of the sponge geometry and measured dimensions.

@@

Finally, it should also be noted that the sponge is reinforced by external ridges that extend perpendicular to the surface of the cylinder and spiral the cage at an angle of $45^\text{o}$. However, in this paper we do not report the effects of these ridges on it's structural performance.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{SFig1.png}
    \caption{\textbf{Hexactinellid sponge \textit{Euplectella aspergillum.}} (a)-(b) Full-frame photo of sponge.  (c)  Close up microscope image of the sponge. (d) Comparison between the idealized model (green and blue lines) and the sponge structure. (e) Unit cell of the idealized model.}
    \label{Sponge}
\end{figure}

\section{Designs Considered}
In this study, we consider four different lattice configurations using in two different elements cross-sections constrained to deform in an in-plane setting only. In an effort to provide a fair performance compassion between the different designs all of the respective lattice configurations, the total amount of material (i.e. volume or mass) is conserved. Moreover,  we consider a fixed volume ratio between non-diagonal and diagonal elements, chosen to match the sponge geometry. In the subsequent sections, we describe in detail the unit cells for the four different designs (Designs A-D).

\subsection{Cylindrical Cross-section}
For this cross-section type we stay consistent with the geometry found within the biological species. Although, this geometry is difficult to realize and validate experimentally, the result for this geometry pose great value to the understanding of the biological species.

\subsubsection{Design A}
Design A is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system (see \cref{DesignA}). If we assume that the horizontal and vertical struts have length $L$ and circular cross section  with diameters $D_{A,nd}$, their volume and projected-area in the unit cell are given by 
\begin{equation}\label{V1}
V_{A,nd}=8 L \left(\pi\frac{D^2_{A,nd}}{4}\right)=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}\label{A1}
A_{A,nd}=8 L D_{A,nd},
\end{equation}
respectively. Note that in this study we use 
\begin{equation}
\frac{D_{A,nd}}{L}=0.1,
\end{equation}
since this is the aspect ratio measured for the sponges (see \cref{sec:params}).

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\linewidth]{SFig2.png}
    \caption{{\bf Unit cell for Design A.} This design is inspired by the sponge structure and consists of a square grid  reinforced by a double diagonal system. The horizontal and vertical struts have length $L$ and circular cross section with diameter $D_{A,nd}$ and, as with the sponge, we assume $D_{A,nd}/L=0.1$. The diagonal elements have a circular cross section  with diameter $D_{A,d}=2 D_{A,nd}$.}
    \label{DesignA}
\end{figure}

\subsubsection{Design B}
Design B is similar to the sponge design (Design A) and is likewise characterized by an alternation of open and closed cells (\cref{DesignB}). However, instead of having two diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. It follows that the non-diagonal and diagonal volumes are given by
\begin{equation}
V_{B,nd}=V_{A,nd}=2\pi L D_{B,nd}^2
\end{equation}
and
\begin{equation}
V_{B,d}=2\sqrt{8} L \left(\pi \frac{{D}_{B,d}^2}{4}\right),
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2}, as well as the above volumes, we  obtain 
\begin{equation}
{{D}_{B,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{B,d}}{{D}_{B,nd}}=\frac{1}{\sqrt{2}}.
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\linewidth]{SFig3.png}
    \caption{{\bf Unit cell for Design B.} This design is still characterized by an alternation of open and closed cells. However, instead of having two  diagonals offset from the nodes, here we only have one diagonal passing through the nodes and crossing every other cell. The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{B,nd}$. The diagonal elements have a circular cross section  with diameter $D_{B,d}={D_{B,nd}}/{\sqrt{2}}$.}
    \label{DesignB}
\end{figure}

\subsubsection{Design C}
Design C is inspired by the town lattice truss design introduced by architect Ithiel Town in 1820 \citep{waddell1916} and consists of every cell being reinforced by diagonal trusses passing through the nodes (see \cref{DesignC}). For this design, the non-diagonal and diagonal volumes of the unit cell are given by:
\begin{equation}
V_{C,nd}=V_{A,nd}=2 L \pi D^2_{A,nd}
\end{equation}
and
\begin{equation}
V_{C,d}=V_{A,d}=2\sqrt{2}L \pi D^2_{A,d},
\end{equation}
respectively.
Using the constraints provided by \cref{con1} and \cref{con2} we  obtain 
\begin{equation}
{{D}_{C,nd}}={{D}_{A,nd}}
\end{equation}
and
\begin{equation}
\frac{{D}_{C,d}}{{D}_{C,nd}}=\frac{1}{2}.
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\linewidth]{SFig4.png}
    \caption{{\bf Unit cell for Design C.} This design consists of a square grid with all cells being reinforced by diagonal trusses passing through the nodes.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{C,nd}$. The diagonal elements have a circular cross section  with diameter $D_{C,d}={D_{C,nd}}/{2}$.}
    \label{DesignC}
\end{figure}

\subsubsection{Design D} 
Design D comprises only the square grid without diagonal reinforcement (\cref{DesignD}) and is well known to be unstable and very limited in resisting shear forces \citep{gibson1999,deshpande2001}. As such, for this design we allocate the total material volume to the non-diagonal elements. Since
\begin{equation}
V_{D,T}=V_{D,nd}=V_{A,nd}=2\pi L D_{D,nd}^2,
\end{equation}
using the constraint provided by \cref{con1} we obtain
\begin{equation}
{D_{D,nd}}={D_{A,nd}}\sqrt{1+\frac{\sqrt{2}}{4}}. 
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\linewidth]{SFig5.png}
    \caption{{\bf Unit cell for Design D.} This design consists of a square grid without diagonal reinforcement.  The horizontal and vertical struts have length $L$ and a circular cross section with diameter $D_{D,nd}$.}
    \label{DesignD}
\end{figure}

\subsection{Cylindrical Cross-section Results}
\mf{add results and figures here}
Include graphs showing results for cylindrical cross-section, as can be found in the main paper.

\subsection{Polygonal Cross-section}
In order to allow for fabrication and testing of plane strain structures and to avoid out of plane buckling, we converted the biological observations of the cylindrical cross-section into a polygonal cross-section. For this new cross-section type, we assume the structure to be sufficiently deep in plane in comparison the structural length scale (plane strain assumption) to avoid out of plane behavior. The cross-section is scaled accordingly with mass while maintaining the depth across all elements constant.

To the determine the minimum cross-section depth a parametric simulation on the necessary depth was run for each structure. The overall minimum depth for all geometries was chosen an a factor of safety of 1.5 was chosen for the finial depth.

\subsubsection{Design A}
From $A_{nd}/A_d\approx 1.4$ we know that the volume is the same given that the in-plane dimension is constant, namely $V_{nd}/V_d\approx 1.4$ . Assuming that the in-plane thickness is given by $t$, we can derive the following volumes for the diagonal and non-diagonal, respectively:
\begin{equation}
	V_{A,nd}=8LD_{A,nd}t
\end{equation}
\begin{equation}
	V_{A,d}=8\sqrt{2}LD_{A,d}t
\end{equation}
Let us enter this into our biological observation, which yields:
\begin{equation}
	\frac{V_{A,nd}}{V_{A,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{A,nd}t}{8\sqrt{2}LD_{A,d}t}
\end{equation}
Therefore, we can simplify this relationship as
\begin{equation}
	D_{A,nd}=2D_{A,d}
\end{equation}

\subsubsection{Design B}
For this design the volume of non-diagonal elements remain the same, namely,
\begin{equation}
		V_{B,nd}=8LD_{B,nd}t.
\end{equation}
However, the volume for the diagonal elements will is different as a result the change in total length, namely,
\begin{equation}
		V_{B,d}=4\sqrt{2}LD_{A,d}t.
\end{equation}
In in effort to maintain a constant volume ration between the diagonally reinforced designs, we use the same volume ratio as per observation
\begin{equation}
		\frac{V_{B,nd}}{V_{B,d}}\approx 1.4\approx\sqrt{2}=\frac{8LD_{B,nd}t}{4\sqrt{2}LD_{B,d}t}
\end{equation}
which simplifies to
\begin{equation}
	D_{B,nd}=D_{B,d}
\end{equation}
and
\begin{equation}
	D_{B,nd}=D_{A,nd}
\end{equation}

\subsubsection{Design D}
For this design, because it does not contain any diagonal elements, we must allocate the total volume of the structure fully on the non-diagonal elements. Thus we can formulate the total volume of Design A as
\begin{equation}
	V_{A,nd}+V_{A,d}=(D_{A,nd}+\sqrt{2}D_{A,d})8Lt
\end{equation}
Using the relationship between the non-diagonal thickness and diagonal thickness, we obtain
\begin{equation}
V_{A,nd}+V_{A,d}=\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}
\end{equation}

We can build a relationship between Design A and Design D through the total volume, such that the total volume of Design A is the same as the total volume of Design D
\begin{equation}
\left(1+\frac{1}{\sqrt{2}}\right)8LtD_{A,nd}=8LtD_{D,nd}
\end{equation}
which yields a relationship between the thicknesses of the non diagonal elements as
\begin{equation}
	D_{D,nd}=\left(1+\frac{1}{\sqrt{2}}\right)D_{A,nd}
\end{equation}

\section{Domain Size Analysis}
In this section we validate the RVE domain size by checking the response of larger domain sizes on the critical buckling value. This ensures that there does not exist another critical buckling mode that is limited by our choice of RVE.

\mf{Write this section}

\section{Optimization Analysis Description}
For the optimization portion of the work presented in the main article we used a stochastic, derivative-free evolutionary method called Covariance Matrix Adaptation Evolution Strategy (CMA-ES) implementation in Python \citep{hansen2003}. The algorithm input variables are: the structure's mass ratio $\lambda$ (keeping the total structure volume constant, as well as distributing the mass evenly through all diagonal and non-diagonal elements respectively) and the separation of the even numbered diagonal elements from the non-diagonal junction. Separate instances of the optimization algorithm are analyzed for different numbers of diagonal reinforcements. For simulations with odd number of diagonal reinforcements, only an even number of diagonals are separated while keeping one diagonal going through the non-diagonal junction in order to ensure geometry symmetry.

For the optimization results presented in the main article, the critical buckling load is the target response maximized. However, a an equivalent analysis is performed to maximize the critical buckling strain as the target response. This analysis is presented in \cref{BucklingOptimization}. Similarly, the same analysis was performed on the elastic stiffness of the structure and yields a trivial result leading to maximizing the allocation of mass to the non-diagonal elements through the mass ratio.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\linewidth]{SFig6.png}
    \caption{Caption}
    \label{BucklingOptimization}
\end{figure}

\section{Parameter Exploration}
\mf{Write this section}
Include the graphs showing critical buckling strain as a function of parameters.
